---
title: "DifferentialAnalysisOfEnhancerSites"
author: "David James"
date: "14/02/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This markdown document performs the main functions 

## Load libraries

```{r libraries}
# Libraries
library(DiffBind)
source("/data/DaveJames/RScripts/createDiffBindTable.R")
library(parallel)
library(rtracklayer)
library(stringr)
library(ChIPpeakAnno)
#source("/data/DaveJames/RScripts/DBA_Helper_Functions.r")

```

```{r}
## Prefix is used to label output files
prefix = "HGSC_Cl_1FC_Cutoff"
outDr = "/data/DaveJames/OC_Epigenetics_Paper/ScriptsForGitHub/testScripts/"
```

## 1. Load GeneHancer database into GRanges object
To get the annotated enhancer locations for differential

```{r genhancer, echo=FALSE}
gh <- read.table("/data/DaveJames/OC_Epigenetics_Paper/genehancer.gff", stringsAsFactors = F, header = T, sep = "\t")
ghGr <- toGRanges(gh, format = "GFF")
ghGrEn <- ghGr[(ghGr$feature.name == "Enhancer" | ghGr$feature.name == "Promoter/Enhancer") ]
```

## 2. Load Required files for comparisons

1) 
```{r pressure, echo=FALSE}

# Upload h3K27ac HGSC cell line data ================================================================================================== #
# tpDr_Celllines <- "/data/ChIPSeqAnalysis/Experiments/OC_Lines_H3K27Ac_GSE156275/results_HGSC_GSE156275/"
# fileIDOC_Celllines <- c("SRR12454735_1", "SRR12454736_1", "SRR12454745", "SRR12454746", "SRR12454748", "SRR12454749")
# bamIdExpInput_Celllines <- c("SRR12454737_1", "SRR12454737_1", "SRR12454747", "SRR12454747", "SRR12454750", "SRR12454750")
# tissue_Celllines <- c("UWB1_289", "UWB1_289", "OAW42", "OAW42", "Kura", "Kura")
# rep_Celllines <- c(1, 2, 1, 2, 1, 2)
# 
# 
# # Upload h3k27ac HGSC primary data ==================================================================================================== #
# tpDrOC_Primary <- "/data/ChIPSeqAnalysis/Experiments/HGSC_H3K27ac/results_HGSC_H3K27ac1/"
# fileIDOC_Primary <- c("SRR8030178.sra", "SRR8030179.sra", "SRR8030180.sra", "SRR8030181.sra", "SRR8030182.sra")
# bamIdExpInput_Primary <- c("SRR8030198.sra", "SRR8030199.sra", "SRR8030200.sra", "SRR8030201.sra", "SRR8030202.sra")
# tissue_Primary <- rep("HGSCPrimary", 5)
# rep_Primary  <- c(1, 2, 3, 4, 5)

# Upload h3K27ac ClearCell cell line data ================================================================================================== #
tpDr_Celllines <- "/data/ChIPSeqAnalysis/Experiments/OC_Lines_H3K27Ac_GSE156275/results_Clearcell_GSE156275//"
fileIDOC_Celllines <- c("SRR12454754", "SRR12454755", "SRR12454728_1", "SRR12454757_1")
bamIdExpInput_Celllines <- c("SRR12454756", "SRR12454756", "SRR12454729_1", "SRR12454723_1")
tissue_Celllines <- c("JHOC5_1", "JHOC5_2", "RMGII_1", "ES2_1")
rep_Celllines <- c(1, 2, 1, 1)

# Upload h3k27ac ClearCell primary data ==================================================================================================== #
tpDrOC_Primary <- "/data/ChIPSeqAnalysis/Experiments/Clearcell_H3K27ac//results_OC_lines_H3K27AC_ClearCell//"
fileIDOC_Primary <- c("SRR8030168.sra", "SRR8030169.sra", "SRR8030170.sra", "SRR8030171.sra", "SRR8030172.sra")
bamIdExpInput_Primary <- c("SRR8030188.sra", "SRR8030189.sra", "SRR8030190.sra", "SRR8030191.sra", "SRR8030192.sra")
tissue_Primary <- rep("ClearCellPrimary", 5)
rep_Primary  <- c(1, 2, 3, 4, 5)

# Upload h3k27ac iOSE  ================================================================================================================ #
tpDrHealthy <- "/data/ChIPSeqAnalysis/Experiments/HealthyOvary_Lines_H3K27ac_GSE68104/results_HealthyOv_GSE68104/"
fileIDHealthy <- c("SRR1983941", "SRR1983942", "SRR1983951", "SRR1983952")
bamIdHealthyInput <- c("SRR1983945", "SRR1983945", "SRR1983955", "SRR1983956")
hlthySmpId <- c("IOE4_1", "IOE4_2", "IOE11_1", "IOE11_2")
htlyRep <- c(1, 2, 1, 2)
htlyTissue <- c("IOE4", "IOE4", "IOE11", "IOE11")

# Create sample tables ================================================================================================================== #
# Cell line sample table
tissue_Healthy_Celllines <- c(htlyTissue, tissue_Celllines)
rep_Healthy_Celllines <- c(htlyRep, rep_Celllines)
sampleID_Healthy_Celllines <- paste(tissue_Healthy_Celllines, rep_Healthy_Celllines, sep = "_")

sampleTableCelllines <- createDiffBindSampleTable(tpDrExp = tpDr_Celllines, tpDrHealthy = tpDrHealthy, sampleID = sampleID_Healthy_Celllines, 
                                                       fileIDExp = fileIDOC_Celllines, fileIDHealthy = fileIDHealthy, bamFiles = T,
                                                       bamFilesHealthInput = bamIdHealthyInput, bamFilesExpInput = bamIdExpInput_Celllines, 
                                                       replicate = rep_Healthy_Celllines, tissue = tissue_Healthy_Celllines)
# Primary sample table
tissue_Healthy_Primaries <- c(htlyTissue, tissue_Primary)
rep_Healthy_Primaries <- c(htlyRep, rep_Primary)
sampleID_Healthy_Primaries <- paste(tissue_Healthy_Primaries, rep_Healthy_Primaries, sep = "_") 

sampleTablePrimaries <- createDiffBindSampleTable(tpDrExp = tpDrOC_Primary, tpDrHealthy = tpDrHealthy, sampleID = sampleID_Healthy_Primaries, 
                                                       fileIDExp = fileIDOC_Primary, fileIDHealthy = fileIDHealthy, bamFiles = T,
                                                       bamFilesHealthInput = bamIdHealthyInput, bamFilesExpInput = bamIdExpInput_Primary, 
                                                       replicate = rep_Healthy_Primaries, tissue = tissue_Healthy_Primaries)

```

## 3. Diffbind Analysis
```{r}
## Diffbind steps for cell lines
# 1 Create cell line sample sheet
dba_Cl <- dba(sampleSheet = sampleTableCelllines)
# 2 Get the cell line counts
dba_Cl_Cnt <- dba.count(dba_Cl, bParallel = TRUE, peaks = ghGrEn, summits = FALSE)
# 3 Normalize the cell line data
dba_Cl_NormDESeq2 <- dba.normalize(dba_Cl_Cnt,  method = DBA_DESEQ2)
# 4 Get the cell line contrast
dba_Cl_NormDESeq2_Contrast <- dba.contrast(dba_Cl_NormDESeq2, reorderMeta = list(Condition="OC"))
# 5 Get the diffrential analysis from the cell lines
dba_Cl_NormDESeq2_DA <- dba.analyze(dba_Cl_NormDESeq2_Contrast, bParallel = TRUE, )

## Diffbind steps for primaries
# 1 Create primaries sample sheet
dba_Prm <- dba(sampleSheet = sampleTablePrimaries)
# 2 Get the primaries counts
dba_Prm_Cnt <- dba.count(dba_Prm, bParallel = TRUE, peaks = ghGrEn, summits = FALSE)
# 3 Normalize the primaries data
dba_Prm_NormDESeq2 <- dba.normalize(dba_Prm_Cnt,  method = DBA_DESEQ2)
# 4 Get the primaries contrast
dba_Prm_NormDESeq2_Contrast <- dba.contrast(dba_Prm_NormDESeq2, reorderMeta = list(Condition="OC"))
# 5 Get the diffrential analysis from the primariess
dba_Prm_NormDESeq2_DA <- dba.analyze(dba_Prm_NormDESeq2_Contrast, bParallel = TRUE, )

```

## 4. Get up/down regulated H3K27ac signal in enhancer sites
```{r}
# Fold change threshold  
fcThres <- 1

# Cell lines
# 1 Get the cell line DE results
dba_Cl_NormDESeq2_DA.report <- dba.report(dba_Cl_NormDESeq2_DA)
# 2 Get the cell line up regulated peaks
up_Cl <- dba_Cl_NormDESeq2_DA.report[dba_Cl_NormDESeq2_DA.report$Fold > fcThres]
# 3 Get the cell line down regulated peaks
dwn_Cl <- dba_Cl_NormDESeq2_DA.report[dba_Cl_NormDESeq2_DA.report$Fold < -fcThres]

# Primaries
# 1 Get the primary DE results
dba_Prm_NormDESeq2_DA.report <- dba.report(dba_Prm_NormDESeq2_DA)
# 2 Get the primary up regulated peaks
up_Prm <- dba_Prm_NormDESeq2_DA.report[dba_Prm_NormDESeq2_DA.report$Fold > fcThres]
# 3 Get the primary down regulated peaks
dwn_Prm <- dba_Prm_NormDESeq2_DA.report[dba_Prm_NormDESeq2_DA.report$Fold < -fcThres]

```

## 5. Get the upregulated peaks 
```{r}

# Get the genehancer peaks which overlap with upregulated cell line peaks
up_Cl_EnhancerIdx <- findOverlaps(ghGrEn, up_Cl)
up_Cl_Enhancer <- ghGrEn[unique(up_Cl_EnhancerIdx@from)]

# Get the genehancer peaks which overlap with upregulated primaries peaks
up_Prm_EnhancerIdx <- findOverlaps(ghGrEn, up_Prm)
up_Prm_Enhancer <- ghGrEn[unique(up_Prm_EnhancerIdx@from)]

# Get the back ground peaks in cell lines
bk_Cl_GR <- makeGRangesFromDataFrame(dba_Cl_NormDESeq2_DA$peaks[[1]])
bk_Cl_EnhancerIdx <- findOverlaps(ghGrEn, bk_Cl_GR)
bk_Cl_Enhancer <- ghGrEn[unique(bk_Cl_EnhancerIdx@from)]

# Get the back ground peaks in primaries
bk_Prm_GR <- makeGRangesFromDataFrame(dba_Prm_NormDESeq2_DA$peaks[[1]])
bk_Prm_EnhancerIdx <- findOverlaps(ghGrEn, bk_Prm_GR)
bk_Prm_Enhancer <- ghGrEn[unique(bk_Prm_EnhancerIdx@from)]

```

## 6. Create final overlapping peak sets and write to MAGIC input file
```{r}

# Get the overlaps between cell lines and primaries
# Back ground
bk_Ovlp_Cl_Prm_EnhancerIdx <- findOverlaps(bk_Cl_Enhancer, bk_Prm_Enhancer)
bk_Ovlp_Cl_Prm_Enhancer <- bk_Cl_Enhancer[unique(bk_Ovlp_Cl_Prm_EnhancerIdx@from)]

# foreground
up_Ovlp_Cl_Prm_EnhancerIdx <- findOverlaps(up_Cl_Enhancer, up_Prm_Enhancer)
up_Ovlp_Cl_Prm_Enhancer <- up_Cl_Enhancer[unique(up_Ovlp_Cl_Prm_EnhancerIdx@from)]
# Write the input file for MAGIC

magicEnh <- data.frame(paste0("GENEHANCER_", str_remove(bk_Ovlp_Cl_Prm_Enhancer@ranges@NAMES, "^X0*")), "" )
colnames(magicEnh) <- c("Background", paste0(prefix, "_enhancers"))
magicEnh[1:length(up_Ovlp_Cl_Prm_Enhancer) , 2] <- paste0("GENEHANCER_", str_remove(up_Ovlp_Cl_Prm_Enhancer@ranges@NAMES, "^X0*"))
write.table(magicEnh, paste0(outDr, prefix, ".txt"), row.names = F, quote = F, sep = "\t")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
